---
title: "R Package Updates"
author: "Jiakun Lin"
date: "`r Sys.Date()`"
output:
  # html_document:
  #   css: styles.css
  slidy_presentation:
    paged: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
load("C:/Users/Kevin/OneDrive - University of Toronto/Summer_Job/Github_IWH/Package6 - BackUp - D matrix+Seperation/IPWMethod/.RData")
library(IPWMethod)
library(tidyverse)
```

## Overview

- Updates to the package  
  - Probability samples ordered by size or as given for precalibration and estimation
  - Survey design of probability samples for D matrix computation

- Demonstrations  

- Questions  


  



## User interface
```{r IPWM, eval=FALSE}
IPWM(
  sc,  
  sp,  
  y,  
  weight,  
  p_formula,  
  method,  
  zcol,  
  precali,  
  
  sp_order,    
  # For multi-reference only
  # If "size", reference samples are reordered by decreasing sample size (default)
  # If "given", preserves the input order
  # The order is used for both precalibration and coefficients estimation
  
  sp_design  
  # Sampling design information for computing the design-based variance component D from the probability sample(s)
)
```


## sp_design
Four design types are available for user:  
```{r sp_design, eval=FALSE}


sp_design = list(type = "poisson") # Poisson (default)

sp_design = list(type = "pps")     # PPS without replacement 

sp_design = list(type = "psu", ids = "psu", strata = "strata", fpc = NULL, nest = TRUE)
# Stratified multistage sampling (PSU-based design)

sp_design = list(type = "replicate", weights = "w", repweights = ..., rep_type = "bootstrap")
# Replicate-weight design (e.g., Statistics Canada bootstrap replicate weights)

```


## PSU-based design
```{r PSU design, eval=FALSE}

design = list(
  type   = "psu",
  ids    = "psu",      # PSU identifier column in sp
  strata = "strata",   # stratum identifier column in sp
  fpc    = NULL,       # optional finite population correction 
  nest   = TRUE        # PSU ids are nested within strata
)

```
keep fpc? expose any additional parameters from the survey package?  

For lonely PSU problem, let user choose handle rule or use "adjust" internally?  
options(survey.lonely.psu = "adjust/certainty/remove")

## Replicate-weight design
```{r replicate design, eval=FALSE}

design = list(
  type       = "replicate",  
  weights    = "w",                     # main weight column name in sp
  repweights = c("rw1","rw2","rw3",...) # replicate weight column names in sp
  rep_type   = "bootstrap"              # only "bootstrap" is currently implemented
           )

```
add other replicate methods (BRR, JK)?


## Example: sp1 with psu/strata infomation
1240 subjects (complete case) from 2005-2010 NHANES wave  
PSU and strata info is included
```{r sp1, echo = F,class="scrollable"}
str(sp1)

```




## sc
1082 subjects (complete case) from 2000-2004 NHANES wave  
height and weight are self-reported to mimic PLCO  
true weights are kept for comparison
```{r sc, echo = F,class="scrollable"}
str(sc)
```

## 
PSU-based
```{r run psu design, echo = F,class="scrollable"}
IPWM(sc=sc,sp=sp1,y="psa",weight="wts_sp1",method="calibration",
     ~agecat + marital + race + educat + empstat + 
        smoking + comorbidity ,
     sp_design=list(type="psu",ids="psu",strata="strata"))
```

Poisson
```{r run poisson, echo = F,class="scrollable"}
IPWM(sc=sc,sp=sp1,y="psa",weight="wts_sp1",method="calibration",
     ~agecat + marital + race + educat + empstat + 
        smoking + comorbidity,
     sp_design=list(type="poisson"))
```

PPS
```{r run pps, echo = F,class="scrollable"}
IPWM(sc=sc,sp=sp1,y="psa",weight="wts_sp1",method="calibration",
     ~agecat + marital + race + educat + empstat + 
        smoking + comorbidity,
     sp_design=list(type="pps"))
```

## Rust and Rao Replication Methods 

For one replicate \(t\):

- From the \(n_h\) sampled PSUs in stratum \(h\), draw \(m_h\) PSUs with replacement \((n_h - 1) \ge m_h > 0\)

- Let \(r_{hi}^{(t)}\) be the number of times PSU \(i\) (in stratum \(h\)) is selected in replicate \(t\).
  \[
  0 \le r_{hi}^{(t)} \le m_h
  \]
- Replicate weights are constructed by:

\vspace{0.3em}

\[
w_{hij}^{(t)}
=
w_{hij}\left\{
\left(1-\sqrt{m_h(n_h-1)}\right)
+
\sqrt{\frac{m_h}{(n_h-1)}}\cdot \frac{n_h}{m_h}\cdot r_{hi}^{(t)}
\right\}.
\]

**Notation**  
- \(h\): stratum,  \(i\): PSU,  \(j\): unit within PSU  
- \(w_{hij}\): original-sample weight;   \(w_{hij}^{(t)}\): replicate-\(t\) weight


## Generate bootstrap weights in sp1
sp1_bt
```{r sp1_bt, echo = F,class="scrollable"}
str(sp1_bt[,1:20])

c(original_total = base_total,
  rep_total_mean = mean(rep_totals),
  rep_sd         = sd(rep_totals))
```

Should we precalibrate bootstrap weights to make them sum up to the same?

## PSU-based design vs. Bootstrap
Bootstrap
```{r run bootstrap, echo = F,class="scrollable"}
IPWM(sc=sc,sp=sp1_bt,y="psa",weight="wts_sp1",method="calibration",
        p_formula = ~agecat + marital + race + educat + empstat + 
        smoking + comorbidity,
     sp_design=list(type="replicate",weights="wts_sp1",
                    repweights=paste0("bw",1:500),rep_type="bootstrap"))
```
PSU
```{r run psu—design, echo = F,class="scrollable"}
IPWM(sc=sc,sp=sp1,y="psa",weight="wts_sp1",method="calibration",
     p_formula = ~agecat + marital + race + educat + empstat + 
        smoking + comorbidity,
     sp_design=list(type="psu",ids="psu",strata="strata"))
```



## sp2
35293 subjects (complete case) from 1997-2008 NHIS (male 50-75 yrs)  
height and weight are self-reported  
PSU and strata info is included
```{r sp2, echo = F}
str(sp2)
```

##    

PSU-based
```{r run psu design sp2, echo = F,class="scrollable"}
IPWM(sc=sc,sp=sp2,y="psa",weight="wts_sp2",method="calibration",
     p_formula = ~ agecat+marital+race+empstat+smoking+comorbidity+BMI,
     sp_design=list(type="psu",ids="psu_sp2",strata="strata_sp2"))
```

Poisson
```{r run poisson sp2, echo = F,class="scrollable"}
IPWM(sc=sc,sp=sp2,y="psa",weight="wts_sp2",method="calibration",
     p_formula = ~ agecat+marital+race+empstat+smoking+comorbidity+BMI,
     sp_design=list(type="poisson"))
```

PPS
```{r run pps sp2, echo = F,class="scrollable"}
IPWM(sc=sc,sp=sp2,y="psa",weight="wts_sp2",method="calibration",
     p_formula = ~ agecat+marital+race+empstat+smoking+comorbidity+BMI,
     sp_design=list(type="pps"))
```



## Multi-reference case

```{r run multi, echo = F,class="scrollable"}
IPWM(sc=sc,sp=list(sp1,sp2),y="psa",weight=c("wts_sp1","wts_sp2"),
     precali=TRUE,sp_order="size",
     p_formula=list(~agecat + marital + race + educat + empstat + 
        smoking + comorbidity,~BMI),
     sp_design=list(
       list(type="psu",ids="psu",strata="strata",nest=TRUE),
       list(type="psu",ids="psu_sp2",strata="strata_sp2",nest=TRUE)
     ))
```



## Results Summary
PSU mean by true weights: 1.762555

```{r results summary, echo = F,class="scrollable"}
res <- data.frame(
  sp_design = c("sp1-PSU", "sp1-Bootstrap", "sp1-Poisson",
                "sp2-PSU", "multi-PSU", "sp1-PSU+psa", "unweighted"),
  Mean = c(1.747084, 1.747084, 1.747084,
           1.813639, 1.767573, 1.979886, 1.969316),
  SE   = c(0.073208, 0.072865, 0.072106,
           0.103363, 0.096465, 0.074997, 0.006298),
  LCI  = c(1.603595, 1.604269, 1.605756,
           1.611048, 1.578501, 1.832893, 1.813772),
  UCI  = c(1.890572, 1.889898, 1.888411,
           2.016230, 1.956644, 2.126879, 2.124860)
)

mu_true <- 1.762555

res[["RB*100"]] <- round((res$Mean - mu_true) / mu_true * 100 , 2)

print(res, row.names = FALSE)


```
Should we also consider binary outcome?


## Seperation of Function
First function has no outcome related variables: outcome(y) and domain variable (zcol)
```{r seperation format, eval=FALSE}
fit = IPWM_build(
  sc,  
  sp,  
  weight,  
  p_formula,  
  method,  
  precali,  
  sp_order
)
```
Second function take the return object from first function, along with outcome(s) and domain variable

```{r seperation format 2, eval=FALSE}
result = IPWM_estimate(
  fit,
  y,  # one outcome at a time
  zcol,
  sp_design 
)
```
